Volunteers in the Clouds: A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2016-01-10 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
Revise authors
And delete index terms as the instructions in the PDF say. Closes #84.
Starting the cuts
I'll be revising on the fly but please @mariosky, @unintendedbear, @pacastillo and @fergunet check that there is no reference to what has gone to the other paper, missing links, whatever. Refs #86

% Some keywords missing: Cloud, Asynchronous, Pool-Based, Javascript-Based
% Open Source software.
% Lets not forget this issue is about software.

\author{Juan-J.~Merelo, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\Mario Garc\'ia-Valdez, Dept. of Graduate Studies, Instituto Tecnológico de Tijuana, Tijuana, M\'exico\\Pedro A. Castillo, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\Pablo Garc\'ia-S\'anchez, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\Paloma de las Cuevas, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\Nuria Rico, Dept. of Statistics and Operational Research, University of Granada, Granada, Spain%
%\author{Juan-J.~Merelo*$^1$, Mario Garc\'ia-Valdez$^2$, Pedro A. Castillo$^1$, Pablo Garc\'ia-S\'anchez$^1$, P. de las Cuevas*$^1$, Nuria Rico$^3$
%\thanks{Manuscript submitted for review on \today.}%
%\thanks{$^1$Dept. of Computer Architecture and Technology and CITIC University of Granada}%
%\thanks{$^2$Dept. of Graduate Studies at Instituto Tecnol\'ogico de Tijuana}%
%\thanks{$^3$Dept. of Statistics and Operational Research University of Granada}%
%\thanks{E-mail addresses: {\tt jmerelo@ugr.es}, {\tt
%    mario@tectijuana.edu.mx}, {\tt pacv@ugr.es}, {\tt pablogarcia@ugr.es},{\tt nrico@ugr.es}}%
\thanks{Corresponding author: Juan-J.~Merelo (Email: jmerelo@ugr.es).}%
}

@@

\begin{IEEEkeywords}
Evolutionary algorithms, evolutionary computation, distributed computing, internet computing,
social networks, volunteer computing, metacomputing, performance evaluation.
\end{IEEEkeywords}

@@

As far as we know, this paper presents one of the few experiments using % (Paloma) Change to "which" in order to not repeating "that" all the time
computational resources that are as dissimilar as smartphones and
powerful laptops 
or workstation computers in a research center. 
%The methodology used
% (Paloma) I would say "the used methodology"
The algorithms used, as well as the methodology 
for gathering resources will be described next, 
together with the results obtained in this initial setup.
%---------------------------------------------------------------
% Start of content for other paper ----
\section{Description of the framework}
\label{sec:description}

In general, a distributed volunteer-based evolutionary computation
framework based on the browser is simply a client-server system
whose client is, or it can be, embedded in the browser via
JavaScript. Since JavaScript is  the only language that is present
across all browsers, the choice was quite clear. Even so, there might
be some issues with the performance of the language itself, which % (Paloma) I recommend the use of "which" here
 is
why we have made a comparison between JavaScript and other languages
\cite{2015arXiv151101088M}. The results of this work show that JavaScript can be
faster than compiled languages such as Scala and, in any case, it provides
a performance that is comparable to other languages usually employed % (Paloma) Adverb
in evolutionary computation and numerical optimization such as Python. 
% (Paloma) Bit of a too long sentence here. Why not changing the ; for a . ?
Compiled languages such as
Java or C might be faster when % (Paloma) Too many "fact" words. There was another one in the previous sentence.
considering the performance of a single-user, 
but this is more than offset by the computational power of
the spontaneous volunteers we can gather at whim, including people
using their mobile phones or tablets. Together, the performance is several orders of magnitude
higher, which is the objective in this kind of systems.
% (Paloma) This sentence needs re-writing
% "Other compiled languages such as Java or C might be, in fact,
% faster, but the fact that single-user performance is higher (than
% what?) is over the offset. This is because the number of spontaneous
% volunteers we can gather at whim, including people using their
% mobile phones or tablets, is several orders of magnitude higher,
% which is the objective in this kind of systems." (?) 
 JavaScript is
quite a popular language nowadays, providing compilers from a number
of languages, many of them in their family (like CoffeeScript), but
many other outside, so researchers can, in fact, write their fitness
function in Python, Erlang, Scala, or even Java, and then compile it to
JavaScript. For completeness, an experiment comparing the performance of
a JavaScript implementation versus Java and Matlab is included in section
\ref{sec:w2}.

As a volunteer based system, there are some issues worth 
noting \cite{sarmenta2001volunteer,web:BOINC} :\begin{enumerate}
\item Volunteers are anonymous, only the Internet Protocol addresses (IP) sent in 
each request is known.
\item As anonymous entities, volunteers are not accountable. 
If a volunteer misbehaves in some way the provider cannot 
prosecute or discipline the volunteer. In the case of JavaScript,
this issue is aggravated because even if it is obfuscated, source code
and data is sent to clients. 
\item Participants must also trust the application provider, 
regarding invasion of privacy, the integrity of their devices, 
and intended application. 
\end{enumerate}

Several kind of threats are possible in this environment; first, since its
programming interface is open and its code is open source, it is
relatively easy to find vulnerabilities and 
sabotage the system, as indicated in \cite{domingues2007sabotage}, by
crafting a fake request which, for instance, assigns a fake fitness to
a particular chromosome.
Other two different kinds of attacks are also
possible: a denial of service attack as well as penetration
testing. We have approached this threat in a pragmatic way: all
sources, client and server, for the application, are openly released
as free software. 
 We also released as an open access paper the result of the 
 experiments performed \cite{DBLP:journals/corr/GuervosG15}  
 and all data, once anonymized, as open data. This
builds trust between the user and the scientist, which indeed has an
impact on the performance of the system, since there is no need to
include cheating checks or other functions that would degrade it. This
socio-technical approach is, indeed, 
coherent with our whole approach to the study of a distributed volunteer
 system which can only be tackled by approaching it from the
technical as well as the social, as in social media, angle.

In this sense, in this paper we propose several versions of the {\sf NodIO} framework, a
cloud or bare metal based volunteer evolutionary computing system
derived from % (Paloma) In order to not sounding too repetitive, could we use "derived from", for example, here?
 the {\sf NodEO} library, whose architecture has been
developed using JavaScript on the client as well as the server.
All parts of the framework are free and available with a free license
from \url{https://github.com/JJ/splash-volunteer}.

%PABLO: el siguiente párrafo justifica Javascript

Thus, {\sf NodIO} architecture has two tiers:\begin{enumerate}
\item A REST server, that is, a server that includes several {\em
  routes} 
  that can be called for storing and retrieving information (the ``CRUD'' cycle:
  create, request, update, and delete) from the server. 
  A JSON data format is used for the communication between 
  clients and the server. There are two kinds of information:
  {\em problem} based, that is, related to the
  evolutionary algorithm such as {\tt PUT}ing a chromosome in or {\tt
  GET}ing a random chromosome from it, and {\em information} related
  to the performance and state of the experiments. It also performs logging
  duties, but they are basically a very lightweight and high performance
  data storage \cite{jj:idc:lowcost}.
  The server has the capability to
  run a single experiment, storing the chromosomes in a data structure
  that is reset when the solution is found.
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays graphs, some
  additional links, and information on the experiment. This code runs
  an evolutionary algorithm {\em island} starting with a random
  population, then it sends, every 100 generations, the best individual
  back to the server (via a {\tt PUT} request), and requests a random
  individual from the server (via a {\tt GET} request). % (Paloma) Re-writing because there are 3 steps (island creation -> PUT -> GET)
\end{enumerate}

  This is the invariant part of the framework, but other than that,
  the algorithm can be run in many different ways: 
  stopping when a solution is found or, as we have done in this paper,
  using Web Workers. In fact, since it is a pool-based system such as
  the one described in \cite{LNCS86720702}, any kind of client that
  calls the application programming interface (API) can be used, % (Paloma) I thought that a client implements an API, or uses an API, or even works with an API. But never heard of a client which "works" an API. Is this correct?
  written in any kind of language. However, in this paper we are going
  to focus on the dynamics and measurements of the volunteer part of
  the framework.

The figure \ref{fig:system} describes the general system architecture and
algorithm behavior. Different web technologies, such as JQuery or {\tt
  Chart.js} have
been used to build the user interface elements of the framework.
\begin{figure}[!t]
\centering
\includegraphics[width=4in]{system.pdf}
\caption{Description of the proposed framework. Clients execute a JavaScript EA
  in the browser, which, every 100 generations, sends the best
  individual and receives a random one from the server.}
\label{fig:system}
\end{figure}

The researcher only has to change a function, which can be written in
JavaScript or in any other language with cross-compilation to
JavaScript \cite{web:compilersjs} to solve a different
problem. In this case the classical Trap function \cite{Ackley1987} has been
used as well as the Rastrigin's floating-point optimization
problem. JavaScript is a functional language and declaring a different 
function and handing at the creation of the algorithm object, called
{\tt Classic}, is the only change needed to work with a different
problem. Let us see how this system addresses the different challenges
outlined in the State of the Art (Section \ref{sec:soa}):
\begin{itemize}
\item {\em Scalability} is provided via the use of a lightweight and
  high-performance, single-threaded, server based in Node.js and
  Express.js. Although this single server is a bottleneck since it
  will eventually saturate, the fact that it runs as a non-blocking single thread
  allows the service of many requests. In fact, a limit in the
  number of simultaneous requests will be reached, but so far it has
  not been found, unlike what we found in our previous systems, DCoR,
  \cite{gecco07:workshop:dcor}, which had a low scalability. 
\item The framework is {\em heterogeneous} since it does not need any
  performance, operating system, or even browser requirement as long
  as JavaScript is enabled: anyone
  visiting the page, even from mobile devices, can load the algorithm.
  In case the browser does not support HTML5 Web workers (e.g., 
  Windows Phone or some version of Android) or their use
  has been disabled, a basic version of {\sf NodIO} can also be used.
\item {\em Fault tolerance} is always an issue, and in this case, the
  single point of failure would be the server: the system, as such,
  would break down if the server fails. However, the individual
  islands in every browser would continue running, and having access
  to just one of them would allow the local algorithm to proceed. In
  fact, the island does not need the server to run: it runs locally if
  needed, with the only exception that it is obviously unable to
  communicate with the rest of the islands.
\item {\em Adaptiveness} is achieved simply through the autonomous
  operation of every individual island without any synchronization
  mechanism. The islands in the system are, in fact, unaware of each
  other, communicating only through the server. This architecture 
  has been tested in other high demand systems with success, but 
  in the case of {\sf NodIO}, additional experiments are needed to asses 
  the scalability of the communication system. % (Paloma) Is it ok if we talk about scalability when we were talking about adaptiveness? Scalability was also described in the first point.
  %(Mario) Scalability of the communication system in particular?
\item Since the algorithm runs in the browser, user's {\em safety} is
  achieved through its sandbox mechanisms. The user is thus assured
  that there is no unsafe access either to their local files or even
  to more resources than the browser should be allotted.
\item Running the algorithm is just a matter of loading the page,
  which makes the operation totally {\em anonymous}. For the same
  reason, {\em ease of use} is optimal, being as easy as simply
  clicking on an URL, available to anyone with access to a browser.
\item {\em Reasonable performance} is not ensured. In fact, we should
  make sure that there is a reasonable amount of clients over which
  the performance achieved is better than what you would obtain in
  your own desktop system. If this is not the case, it is a pointless
  academic exercise.
\end{itemize}

There are many different ways to validate a framework that intends to
address these challenges. Some of them are included by design: it is
anonymous, since no accounts are needed to participate in the
experiment, it is safe for the user, since we are using the browser
black box, and it is also heterogeneous. We have designed the
experiments so that they show adaptability to different types of
experiments and clients, scalability with the number of clients and
the problem size, and measured performance by comparing it with
a baseline configuration, with which, as indicated above, is
compatible and can be used in a complementary way. Fault tolerance
will be experimented by dropping the server and checking the continuity
of the experiments.

All these points will be established in the next two sections,
starting with the baseline, single worker system next.

%---------------------------------------------------------------
\section{Baseline Experiments} %Cambié el nombre
\label{sec:experiments}

\begin{figure}[!t]
\centering
\includegraphics[width=10cm]{baseline-times.png}
\caption{Comparison of times to solution (labeled as {\bf times}) for the baseline system, using a {\tt
    Node.js} client, for different population sizes. Only
runs in which solution has been found are used for this graph.}
% Plotted with: ~/Code/splash-volunteer/data/plot-baseline.R
\label{fig:baseline}
\end{figure}
First, we establish the baseline performance by running fifty runs of the
evolutionary algorithm in a desktop client written using {\sf NodEO}
\cite{nodeo2014}, the basic JavaScript library we have used as base to
build {\sf NodIO}, the framework. This experiment tries to 
find the solution to the 40-trap function with parameters $l=4, a=1,
b=2, z=3$ and a population size equal to 512. The algorithm in a
single island was run until the solution (a string with all
ones) was found or five million evaluations had been performed. It
took around a minute, on average, that is $68.9694$ seconds, to
perform these runs. In this experiment only 33, that is, 66\% were
successful. The experiments were repeated for population $p=1024$,
with a success rate upgraded to 100\% and an {\em average} duration of $3.46$
seconds. Results for these two experiments
are shown in Figure \ref{fig:baseline}. The baseline hardware system has these
characteristics: 
\begin{verbatim}
Linux penny 3.13.0-34-generic #60-Ubuntu SMP Wed Aug 13 15:45:27 UTC
2014 x86_64 x86_64 x86_64 GNU/Linux
\end{verbatim}
with a 4-core {\tt Intel(R) Core(TM) i7-4770 CPU @ 3.40GHz}.

% end spin-off ------------------------------------------------------

>>>>

\author{Juan-J.~Merelo, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\Mario Garc\'ia-Valdez, Dept. of Graduate Studies, Instituto Tecnológico de Tijuana, Tijuana, M\'exico\\Pedro A. Castillo, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\Pablo Garc\'ia-S\'anchez, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\Paloma de las Cuevas, Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain\\Nuria Rico, Dept. of Statistics and Operational Research, University of Granada, Granada, Spain%

\thanks{Corresponding author: Juan-J.~Merelo (Email: jmerelo@ugr.es).}%
}

@@

@@

As far as we know, this paper presents one of the few experiments using % (Paloma) Change to "which" in order to not repeating "that" all the time
computational resources that are as dissimilar as smartphones and
powerful laptops 
or workstation computers in a research center. 
%The methodology used
% (Paloma) I would say "the used methodology"
The algorithms used, as well as the methodology 
for gathering resources, will be described next, %change this to
                                %references to the other paper - JJ
together with the results obtained in this initial setup.

\section{Experimental results}
\label{sec:exp1}
